---
title: "ML Forecast Final"
author: "Tee Wei Hern"
date: "2025-08-13"
output: html_document
---

```{r}
library(readxl) # read xlsx file
library(openxlsx) # export xlsx file
library(glmnet) # lasso
library(tidyverse)
library(caret) # random forest
library(ggplot2)
library(forecast) # arima/random walk
library(zoo) # as.yearqtr()
```

# Cleaning data
```{r}
inflation_data_raw <- read_excel("N:July 2025 Inflation Data.xlsx")
```

```{r}
inflation_data <- inflation_data_raw %>%
  select(!c(headline : supercore10)) %>%
  select(!c(petrol95 : ...51)) %>%
  select(!c(...54 : ...70, "Paste ygap here", "Half rule")) %>%
  rename(date = ...1) %>%
  rename(er = er...18)

inflation_data$winf <- as.numeric(inflation_data$winf)
inflation_data$neer <- as.numeric(inflation_data$neer)
inflation_data$comm <- as.numeric(inflation_data$comm)
inflation_data$er <- as.numeric(inflation_data$er)
inflation_data$ygap <- as.numeric(inflation_data$ygap)
inflation_data$wage <- as.numeric(inflation_data$wage)
inflation_data$prod <- as.numeric(inflation_data$prod)

# transforming data to qoq
inflation_data <- inflation_data %>%
  mutate(core_qoq = 100 * (log(inflation_data$core) - log(lag(inflation_data$core))),
         winf_qoq = 100 * (log(inflation_data$winf) - log(lag(inflation_data$winf))),
         neer_qoq = 100 * (log(inflation_data$neer) - log(lag(inflation_data$neer))),
         comm_qoq = 100 * (log(inflation_data$comm) - log(lag(inflation_data$comm))),
         er_qoq = 100 * (log(inflation_data$er) - log(lag(inflation_data$er))),
         wage_qoq = 100 * (log(inflation_data$wage) - log(lag(inflation_data$wage))),
         prod_qoq = 100 * (log(inflation_data$prod) - log(lag(inflation_data$prod))))

# adding lagged variables
inflation_data <- inflation_data %>%
  mutate(core_lag1 = lag(core_qoq, 1),
         core_lag4 = lag(core_qoq, 4),
         winf_lag1 = lag(winf_qoq, 1),
         winf_lag4 = lag(winf_qoq, 4),
         neer_lag1 = lag(neer_qoq, 1),
         neer_lag4 = lag(neer_qoq, 4),
         comm_lag1 = lag(comm_qoq, 1),
         comm_lag4 = lag(comm_qoq, 4),
         er_lag1 = lag(er_qoq, 1),
         er_lag4 = lag(er_qoq, 4),
         ygap_lag1 = lag(ygap_qoq, 1),
         ygap_lag4 = lag(ygap_qoq, 4),
         wage_lag1 = lag(wage_qoq, 1),
         wage_lag4 = lag(wage_qoq, 4),
         prod_lag1 = lag(prod_qoq, 1),
         prod_lag4 = lag(prod_qoq, 4))
```
```{r}
inflation_data <- inflation_data %>%
  slice(-c(145:154))
```
```{r}
# split training and testing set following chronological order
inf_train <- inflation_data %>%
  slice(c(1:132))

inf_test <- inflation_data %>%
  slice(c(133:137))

inf_forecast <- inflation_data %>%
  slice(c(137:144))
```

# 1 Random Forest
```{r}
# Remove NA values
inf_train <- inflation_data %>% drop_na()

# Parameter tuning
rf_grid <- expand.grid(mtry = c(3, 5, 7, 9), min.node.size = c(10, 15, 20), splitrule = "variance")

# Train parameters, using time series cross validation (TSCV) for data that has temporal order
set.seed(1)
fit_Control <- trainControl(method = "timeslice",
                            initialWindow = 80, # size of initial training
                            horizon = 1, # forecast horizon
                            fixedWindow = TRUE, # training set shifts each iteration(rolling origin)
                            summaryFunction = defaultSummary)

set.seed(1)
RF_model <- train(core_qoq ~ winf_qoq + neer_qoq + comm_qoq + er_qoq + ygap + wage_qoq + prod_qoq + core_lag1 + core_lag4 + winf_lag1 + winf_lag4 + neer_lag1 + neer_lag4 + comm_lag1 + comm_lag4 + er_lag1 + er_lag4 + ygap_lag1 + ygap_lag4 + wage_lag1 + wage_lag4 + prod_lag1 + prod_lag4 + q1 + q2 + q3 + q4, 
                  data = inf_train,
                  method = "ranger",
                  trControl = fit_Control,
                  tuneGrid = rf_grid,
                  num.trees = 300,
                  importance = 'permutation')

# Check variable importance for each predictor
varImp(RF_model)

print(RF_model$results %>% arrange(RMSE))
```
```{r}
# visualize resampling results
ggplot(RF_model, metric = "RMSE")
```
```{r}
fitControl_final <- trainControl(method = "none")
RF_final <- train(core_qoq ~ winf_qoq + neer_qoq + comm_qoq + er_qoq + ygap + wage_qoq + prod_qoq + core_lag1 + core_lag4 + winf_lag1 + winf_lag4 + neer_lag1 + neer_lag4 + comm_lag1 + comm_lag4 + er_lag1 + er_lag4 + ygap_lag1 + ygap_lag4 + wage_lag1 + wage_lag4 + prod_lag1 + prod_lag4 + q1 + q2 + q3 + q4,
                  data = inf_train,
                  method = "ranger",
                  trControl = fitControl_final,
                  tuneGrid = data.frame(mtry = 7, min.node.size = 15, splitrule = "variance"), num.trees = 300, importance = 'permutation')

varImp(RF_final)
```
```{r}
# Perform prediction
RF_predict_train <- predict(RF_final, newdata = inf_train)
RF_predict_test <- predict(RF_final, newdata = inf_test)
```
```{r}
# Creating line plot
results <- data.frame(
  Date = inf_train$date,
  Actual = inf_train$core_qoq,
  Predicted = RF_predict_train)

ggplot(results, aes(x = Date)) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Predicted, color = "Predicted")) +
  labs(title = "Actual vs Predicted Core Inflation (Random Forest)",
       y = "Core Inflation(%)", x = "Date", color = "Legend") +
  theme_minimal()

rmse <- sqrt(mean((inf_train$core_qoq - RF_predict_train)^2))
print(rmse)
```
```{r}
# Repeat with testing data
results_RF <- data.frame(
  Date = inf_test$date,
  Actual = inf_test$core_qoq,
  Predicted = RF_predict_test)

ggplot(results_RF, aes(x = Date)) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Predicted, color = "Predicted")) +
  geom_line(aes(y = Actual, color = "Actual"), size = 3) +
  geom_line(aes(y = Predicted, color = "Predicted"), size = 3) +
  labs(title = "Actual vs Predicted Core Inflation (Random Forest)",
       y = "Core Inflation (%)", x = "Date", color = "Legend") +
  theme_minimal()

rmse <- sqrt(mean((inf_test$core_qoq - RF_predict_test)^2))
print(rmse)
```

```{r}
# Creating rolling window forecast 2025Q2 - 2026Q4
forecast_core_qoq_RF <- function(model, data) {
  n <- nrow(data)
  
  # Store forecasts
  forecasts <- rep(NA, n)
  
  # Loop through each row
  for (i in 1:n) {
    # Only forecast if core_qoq is NA
    if (is.na(data$core_qoq[i])) {
      
      # Prepare predictors for this row
      predictors <- data[i, ] %>%
        select(winf_qoq, neer_qoq, comm_qoq, er_qoq, ygap, wage_qoq, prod_qoq, winf_lag1, winf_lag4, neer_lag1, neer_lag4, comm_lag1, comm_lag4, er_lag1, er_lag4, ygap_lag1, ygap_lag4, wage_lag1, wage_lag4, prod_lag1, prod_lag4, core_lag1, core_lag4, q1, q2, q3, q4)
      
      # Convert to matrix for LASSO/RF
      X <- as.matrix(predictors)
      
      # Forecast using model
      forecast <- predict(model, newdata = X)
      
      # Save forecast
      data$core_qoq[i] <- forecasts
      forecasts[i] <- forecast
      
      # Update lagged values in future rows
      if (i + 1 <= n) {
        data$core_lag1[i + 1] <- forecast
      }
      if (i + 4 <= n) {
        data$core_lag4[i + 4] <- forecast
      }
    }
  }
  return(list(updated_data = data, forecasts = forecasts))
}

forecast_core_qoq_RF(RF_final, inf_forecast)
```
```{r}
results_forecast_RF <- data.frame(Date = inf_forecast$date,
                                  Predicted = forecast_core_qoq_RF(RF_final, inf_forecast)$forecasts)
```

# 2 LASSO
```{r}
# Remove NA values
inf_train <- inf_train %>% drop_na()

y_train <- inf_train$core_qoq
X_train <- as.matrix(inf_train %>%
                       select(winf_qoq, neer_qoq, comm_qoq, er_qoq, ygap, wage_qoq, prod_qoq, winf_lag1, winf_lag4, neer_lag1, neer_lag4, comm_lag1, comm_lag4, er_lag1, er_lag4, ygap_lag1, ygap_lag4, wage_lag1, wage_lag4, prod_lag1, prod_lag4, core_lag1, core_lag4, q1, q2, q3, q4))

y_test <- inf_test$core_qoq
X_test <- as.matrix(inf_test %>%
                       select(winf_qoq, neer_qoq, comm_qoq, er_qoq, ygap, wage_qoq, prod_qoq, winf_lag1, winf_lag4, neer_lag1, neer_lag4, comm_lag1, comm_lag4, er_lag1, er_lag4, ygap_lag1, ygap_lag4, wage_lag1, wage_lag4, prod_lag1, prod_lag4, core_lag1, core_lag4, q1, q2, q3, q4))

# Scaling training and testing set(already scaled since qoq)
X_scaled_train <- X_train
X_scaled_test <- X_test

# Perform time series CV, along with testing best alpha, best lambda, and best horizon
tscv_glmnet_alpha_search <- function(X, y, alpha_seq = seq(0, 1, by = 0.1), 
                                     lambda_seq = 10^seq(-4, 1, length.out = 100),
                                     initial_window = 80, 
                                     horizon) {
  n_obs <- nrow(X)
  n_folds <- n_obs - initial_window
  results <- data.frame(alpha = numeric(), lambda = numeric(), rmse = numeric())
  
  for (alpha_val in alpha_seq) {
    rmse_per_lambda <- numeric(length(lambda_seq))
    
    for(j in seq_along(lambda_seq)) {
      lambda <- lambda_seq[j]
      preds <- c()
      actuals <- c()
      
      for(i in 1:(n_folds - horizon + 1)) {
        idx_train <- 1:(initial_window + i - 1)
        idx_val <- (initial_window + i):(initial_window + i + horizon - 1)
        
        X_t <- X[idx_train, ]
        y_t <- y[idx_train]
        X_val <- X[idx_val, , drop = FALSE]
        y_val <- y[idx_val]
        
        model <- glmnet(X_t, y_t, alpha = alpha_val, lambda = lambda)
        pred <- predict(model, newx = X_val)
        
        preds <- c(preds, pred)
        actuals <- c(actuals, y_val)
      }
      rmse_per_lambda[j] <- sqrt(mean((actuals - preds)^2))
    }
    best_lambda <- lambda_seq[which.min(rmse_per_lambda)]
    best_rmse <- min(rmse_per_lambda)
    
    results <- rbind(results, data.frame(alpha = alpha_val, lambda = best_lambda, rmse = best_rmse))
  }
  
  best_result <- results[which.min(results$rmse), ]
  return(list(best_alpha = best_result$alpha,
              best_lambda = best_result$lambda,
              best_rmse = best_result$rmse,
              all_results = results))
}

# Testing over horizons of 1, 4 (quarterly, annually)
horizons <- c(1, 4)
results_by_horizon <- list()

for (h in horizons) {
  result <- tscv_glmnet_alpha_search(X_scaled_train, y_train, horizon = h)
  results_by_horizon[[paste0("horizon_", h)]] <- result
  print(results_by_horizon[[paste0("horizon_", h)]])
}

# Extract RMSE from each horizon result
rmse_values <- sapply(results_by_horizon, function(res) res$best_rmse)

# Identify the best horizon
best_horizon_name <- names(rmse_values)[which.min(rmse_values)]
best_result <- results_by_horizon[[best_horizon_name]]

# Print the best results
cat("Best Horizon:", best_horizon_name, "\n")
cat("Best Alpha:", best_result$best_alpha, "\n")
cat("Best Lambda:", best_result$best_lambda, "\n")
cat("Best RMSE:", best_result$best_rmse, "\n")
```

```{r}
# Train final model on entire training set with best lambda
final_model <- glmnet(X_scaled_train, y_train, alpha = result$best_alpha, lambda = result$best_lambda)

# Predict on training and testing set
y_pred_train <- predict(final_model, newx = X_scaled_train)
y_pred_test <- predict(final_model, newx = X_scaled_test)
rmse_train <- sqrt(mean((y_train - y_pred_train)^2))
rmse_test <- sqrt(mean((y_test - y_pred_test)^2))
print(rmse_train)
print(rmse_test)
```
```{r}
results <- data.frame(
  Date = inf_train$date,
  Actual = y_train,
  Predicted = as.numeric(y_pred_train))

results <- na.omit(results)

ggplot(results, aes(x = Date)) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Predicted, color = "Predicted")) +
  labs(title = "LASSO: Actual vs Predicted Core Inflation",
       y = "Core Inflation (%)", color = "", x = "Date") +
  theme_minimal()

rmse_train
```
```{r}
results <- data.frame(
  Date = inf_test$date,
  Actual = y_test,
  Predicted = as.numeric(y_pred_test))

results <- na.omit(results)

ggplot(results, aes(x = Date)) +
  geom_point(aes(y = Actual, color = "Actual"), size = 3) +
  geom_point(aes(y = Predicted, color = "Predicted"), size = 3) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Predicted, color = "Predicted")) +
  labs(title = "LASSO: Actual vs Predicted Core Inflation",
       y = "Core Inflation (%)", color = "", x = "Date") +
  theme_minimal()

rmse_test
```

```{r}
# Plotting graph of RMSE against alpha value
ggplot(results$all_results, aes(x = alpha, y = rmse)) +
  geom_point(color = "blue", size = 2) +
  geom_line(color = "blue", linewidth = 1) +
  labs(title = "RMSE vs Alpha (glmnet TSCV)",
       x = "Alpha (0 = Ridge, 1 = LASSO)",
       y = "RMSE") +
  theme_minimal()
```
```{r}
# Extract coefficients from the final model
coef(final_model)
```

```{r}
# Creating rolling window forecast 2025Q2 - 2026Q4
forecast_core_qoq_lasso <- function(model, data) {
  n <- nrow(data)
  
  # Store forecasts
  forecasts <- rep(NA, n)
  
  # Loop through each row
  for (i in 1:n) {
    # Only forecast if core_qoq is NA
    if (is.na(data$core_qoq[i])) {
      
      # Prepare predictors for this row
      predictors <- data[i, ] %>%
        select(winf_qoq, neer_qoq, comm_qoq, er_qoq, ygap, wage_qoq, prod_qoq, winf_lag1, winf_lag4, neer_lag1, neer_lag4, comm_lag1, comm_lag4, er_lag1, er_lag4, ygap_lag1, ygap_lag4, wage_lag1, wage_lag4, prod_lag1, prod_lag4, core_lag1, core_lag4, q1, q2, q3, q4)
      
      # Convert to matrix for LASSO
      X <- as.matrix(predictors)
      
      # Forecast using model
      forecast <- predict(model, newx = X)
      
      # Save forecast
      data$core_qoq[i] <- forecasts
      forecasts[i] <- forecast
      
      # Update lagged values in future rows
      if (i + 1 <= n) {
        data$core_lag1[i + 1] <- forecast
      }
      if (i + 4 <= n) {
        data$core_lag4[i + 4] <- forecast
      }
    }
  }
  return(list(updated_data = data, forecasts = forecasts))
}

forecast_core_qoq_lasso(final_model, inf_forecast)
```
```{r}
results_forecast_lasso <- data.frame(Date = inf_forecast$date,
                                  Predicted = forecast_core_qoq_lasso(final_model, inf_forecast)$forecasts)
```

# 3 ARIMA as non-ML baseline
```{r}
# converting data to time series object
inf_ts <- ts(inflation_data$core_qoq, start = c(1991, 1), frequency = 4)  # quarterly
train <- window(inf_ts, end = c(2023,4))
test <- window(inf_ts, start = c(2024, 1))
```
```{r}
# fit ARIMA model
arima_model <- auto.arima(train)
summary(arima_model)
```
```{r}
# forecast future values
n_test <- length(test)
forecast_arima <- forecast(arima_model, h = 5) # make consistent with testing window
plot(forecast_arima)
lines(test, col = "red")

accuracy(forecast_arima, test)
```

# 4 Compare predictions by backtesting (with NKPC)
```{r}
results_test <- data.frame(
  Date = inf_test$date,
  Actual = inf_test$core_qoq,
  LASSO = as.numeric(y_pred_test),
  RF = RF_predict_test,
  ARIMA = as.numeric(forecast_arima$mean))

library(ggplot2)
ggplot(results_test, aes(x = Date)) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = LASSO, color = "LASSO")) +
  geom_line(aes(y = RF, color = "Random Forest")) +
  geom_line(aes(y = ARIMA, color = "ARIMA"))

# Converting qoq inflation to yoy 
results_train <- data.frame(
  Date = inf_train$date,
  Actual = y_train,
  LASSO = as.numeric(y_pred_train),
  RF = RF_predict_train,
  ARIMA = inf_train$core_qoq) # training same as actual

results_join <- bind_rows(results_train, results_test)

# To get qoq values from 2023 to transform into yoy in 2024
results_join <- results_join %>%
  slice_tail(n = 9)
```

```{r}
# Export results_join to excel to perform function externally (qoq to yoy)
library(openxlsx)

# Only run this once, if data is already written to excel, skip this step
# Create this data on a separate tab / excel
wb <- loadWorkbook("copy file path here")
writeData(wb, sheet = "Sheet 2 (Example)", x = results_join, startCol = 1, startRow = 1)
saveWorkbook(wb, "copy file path here", overwrite = TRUE)
```

```{r}
# Import data from the excel above after performing function to transform yoy
inf_yoy_raw <- read_excel("copy file path here", sheet = "Sheet 2 (Example)")

# Update this according to testing period, or just use inf_yoy_raw$date
yoy <- c("2024Q1", "2024Q2", "2024Q3", "2024Q4", "2025Q1")

inf_yoy <- inf_yoy_raw %>%
  select(c(inf_yoy_actual: inf_yoy_ARIMA)) %>% # data here transformed to yoy
  drop_na()

# Importing NKPC info
NKPC_2023 <- read_excel("path to NKPC values 2023", sheet = "Sheet 2 (Example)")
NKPC_2024 <- read_excel("path to NKPC values 2024", sheet = "Sheet 2 (Example)")
NKPC_2025 <- read_excel("path to NKPC values 2025", sheet = "Sheet 2 (Example)")

# slice NKPC to back-testing period
NKPC_2024 <- NKPC_2024 %>%
  slice_head(n = 5)
```
```{r}
inf_yoy <- data.frame(Period_yoy = yoy, # or get from inf_yoy_raw$date
                      Actual = inf_yoy$inf_yoy_actual,
                      LASSO = inf_yoy$inf_yoy_lasso,
                      RF = inf_yoy$inf_yoy_RF,
                      ARIMA = inf_yoy$inf_yoy_ARIMA,
                      NKPC = NKPC_2024$F_NKPC)

print(inf_yoy)
```
```{r}
ggplot(inf_yoy, aes(x = Period_yoy)) +
  geom_point(aes(y = Actual, color = "Actual"), size = 3) +
  geom_point(aes(y = LASSO, color = "LASSO"), size = 3) +
  geom_point(aes(y = RF, color = "RF"), size = 3) +
  geom_point(aes(y = ARIMA, color = "ARIMA"), size = 3) +
  geom_point(aes(y = NKPC, color = "NKPC"), size = 3) +
  
  geom_line(aes(y = Actual, color = "Actual"), group = 1) +
  geom_line(aes(y = LASSO, color = "LASSO"), group = 1) +
  geom_line(aes(y = RF, color = "RF"), group = 1) +
  geom_line(aes(y = ARIMA, color = "ARIMA"), group = 1) +
  geom_line(aes(y = NKPC, color = "NKPC"), group = 1) +
  labs(title = "Inflation Forecast = Actual vs Predicted Core Inflation",
       y = "Core Inflation", color = "", x = "Period") +
  theme_minimal()

# bar graph yoy (remove ARIMA to compare only ML and NKPC)
inf_yoy_long <- inf_yoy %>%
  pivot_longer(cols = c("Actual", "LASSO", "RF", "NKPC"), 
               names_to = "Type", 
               values_to = "Inflation")

inf_yoy_long$Type <- factor(inf_yoy_long$Type, levels = c("Actual", "LASSO", "RF", "NKPC"))

ggplot(inf_yoy_long, aes(x = Period_yoy, y = Inflation, fill = Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = sprintf("%.2f", Inflation)),
            position = position_dodge(width = 0.9),
            vjust = 1.5) +
  labs(title = "Annual Actual vs Predicted Core Inflation",
       x = "Year", y = "Annual Core Inflation (%)", fill = "") +
  theme_minimal()
```
```{r}
# Calculating RMSE
sqdiff <- inf_yoy_raw %>% # squared difference calculated from excel as well
  select(c(sqdiff_lasso : sqdiff_ARIMA)) %>%
  drop_na()

NKPC_2024_sq_diff <- NKPC_2024 %>%
  select(F_NKPC) %>%
  transmute(sqdiff_NKPC = (F_NKPC - inf_yoy$Actual)^2)

sqdiff_all <- cbind(sqdiff, NKPC_2024_sqdiff)

RMSE_2024 <- sqdiff_all %>%
  slice_head(n = 4) %>%
  summarise(RMSE_lasso = sqrt(mean(sqdiff_lasso)),
            RMSE_RF = sqrt(mean(sqdiff_RF)),
            RMSE_ARIMA = sqrt(mean(sqdiff_ARIMA)),
            RMSE_NKPC = sqrt(mean(sqdiff_NKPC)))

print(RMSE_2024)
```

# 5 Forecasting ahead 2025Q2 to 2026Q
```{r}
# Import excel file which contains forecasted values
inf_yoy_forecast <- read_excel("Copy file path here")
inf_forecast_final <- inf_yoy_forecast %>%
  select(c(Date, inf_yoy_lasso : inf_yoy_ARIMA)) %>%
  slice_tail(n = 7)

NKPC_2025_cut <- NKPC_2025 %>%
  select(F_NKPC) %>%
  slice_tail(n = 7)

inf_forecast_final_NKPC <- cbind(inf_forecast_final, NKPC_2025_cut)

print(inf_forecast_final_NKPC)
```
```{r}
# plotting graph of forecast
yoy <- c("2025Q2", "2025Q3", "2025Q4", "2026Q1", "2026Q2", "2026Q3", "2026Q4")
inf_forecsat_final_NKPC$Period_yoy <- yoy # add new column

ggplot(inf_forecast_final_NKPC, aes(x = Period_yoy)) +
  geom_point(aes(y = inf_yoy_lasso, color = "LASSO"), size = 3) +
  geom_point(aes(y = inf_yoy_RF, color = "RF"), size = 3) +
  geom_point(aes(y = inf_yoy_ARIMA, color = "ARIMA"), size = 3) +
  geom_point(aes(y = F_NKPC, color = "F_NKPC"), size = 3) +
  geom_line(aes(y = inf_yoy_lasso, color = "LASSO"), group = 1) +
  geom_line(aes(y = inf_yoy_RF, color = "RF"), group = 1) +
  geom_line(aes(y = inf_yoy_ARIMA, color = "ARIMA"), group = 1) +
  geom_line(aes(y = F_NKPC, color = "F_NKPC"), group = 1) +
  labs(title - "Inflation Forecast 2025Q2 to 2026Q4",
       y = "Core Inflation (%)", color = "", x = "Period") +
  theme_minimal()

#bar graph
inf_forecast_final_NKPC_bar <- inf_forecast_final_NKPC %>%
  select(!c(inf_yoy_ARIMA)) %>%
  rename(LASSO = inf_yoy_lasso,
         RF = inf_yoy_RF,
         NKPC = F_NKPC)

inf_yoy_forecast_long <- inf_forecast_final_NKPC_bar %>%
  pivot_longer(cols = c("LASSO", "RF", "NKPC"),
               names_to = "Type", values_to = "Inflation")

inf_yoy_forecast_long$Type <- factor(inf_yoy_forecast_long$Type, levels = c("LASSO", "RF", "NKPC"))

ggplot(inf_yoy_forecast_long, aes(x = Period_yoy, y = Inflation, fill = Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = sprintf("%.2f", Inflation)),
            position = position_dodge(width = 0.9),
            vjust = 1.5) +
  labs(title = "Core Inflation Forecast",
       x = "Year", y = "Annual Core Inflation (%)", fill = "") +
  theme_minimal()
```
**Thing to consider: model is trained with log diff qoq. yoy calculation is marginally larger than log diff yoy (adding 4 quarters of log diff qoq), but difference is largely negligible.**
**Moving ahead: including other features not in NKPC to see if it is an important determinant in predicting inflation (KLCI, card spending etc); can attempt to automate code even further.**

# Guide to import updated dataset to train ML models
## Disclaimer: Other than the lines stated later below, there may be occasions where changes have to be done on other lines to standardize imported data sets, hence cleaning data is very important.

Note: Ensure all names are consistent in order for code to run successfully; relevant data cleaning has to be done every time an file(excel) is imported to ensure consistency

Line 21 - Import updated data set. (Core sheet)

Line 24 to 68 - Clean data as needed, convert all determinants to log diff qoq (or can do in excel), then add lags to variables.

Line 69 to 73 - Slice data appropriately for training, can just move one row down if data set is updated quarterly.

## Run random forest and LASSO codes ("#1 and #2")

## 3 ARIMA
Line 419 to 420 - Update period for training and testing window

## 4 Compare predictions by backtesting (with NKPC)
Note: Many of this part may be optional as they are mostly codes to visualize results, but the crucial part is transforming qoq inflation to yoy. This is done by exporting the model predicted values to excel to transform qoq to yoy.

Line 472 to 476 - Export qoq predictions to excel 

Line 481 - Import data set that has already been transformed to yoy
Note: sqdiff (squared difference) between predicted and actual values are also calculated in excel in separate columns.

Line 491 to 493 - Import NKPC data from excel

Line 479 to 507 - Clean all imported data appropriately to obtain desired inf_yoy table containing values of date, actual inflation, and forecast from all models. inf_yoy is also used for creating line and bar graphs (if needed)

## 5 Forecasting ahead 2025Q2 to 2026Q
Line 568 to 571 - Process similar to ##4, but imported data set has forecast values from 2025Q2 to 2026Q4 as well

Line 566 to 580 - Clean data so that only forecast values remain from 2025Q2 to 2026Q4